Laws and institutions shape individual outcomes through complex interactions with
citizens' diverse circumstances, yet how different voting methods navigate this
coupled landscape remains poorly understood. We model collective governance as
optimization on NK fitness landscapes, where shared bits (laws) are updated by
voting while individual bits (personal traits) remain fixed. A cross-dependency
parameter $\alpha$ controls how legislation's effects depend on individual
circumstances. We compare eight standard voting methods and a generalized
scoring family across landscape ruggedness $K \in \{1,\ldots,20\}$ and
$\alpha \in [0,1]$ with 1000 runs per configuration.

We find that the optimal voting method undergoes sharp phase transitions as a
function of landscape complexity: total score voting dominates on smooth
landscapes, a generalized scoring rule with $p=0.35$ at low-to-moderate
ruggedness, Borda count across a wide middle range, and STAR voting at the
highest complexity. A two-parameter empirical formula reduces the
$(K, \alpha)$ plane to a single complexity axis for visualization. Borda count
emerges as uniquely robust, achieving the highest mean fitness and lowest
variance across most of the parameter space.

We further introduce a representative democracy model parameterized by identity
weight $\beta$ and candidate selfishness $p_{\mathrm{self}}$. Representation
reshapes the complexity-dependent structure even under favorable conditions: total
score voting dominates across most regimes, and as identity voting increases,
plurality performs surprisingly well while sophisticated methods lose their
advantage.
