All experiments use $N=50$ bit positions with a voting portion of $0.5$
($N_s = N_t = 25$), a population of $M=100$ individuals, $V=2$ bits voted on per
round, and all individuals sharing identical initial voting bits with unique,
randomly generated individual bits.

\subsection{Experiment 1: $K \times \alpha$ Sweep (Direct Democracy)}

We sweep landscape ruggedness $K \in \{1, 2, \ldots, 20\}$ and cross-dependency
$\alpha \in \{0.00, 0.05, 0.10, \ldots, 1.00\}$ (21 levels) under direct
democracy for all eight standard voting methods. Each configuration is run for
$150 + 50K$ iterations with 1000 independent runs, providing high-resolution
coverage of the full $(K, \alpha)$ plane. For each run we record mean fitness,
fitness variance, minimum and maximum fitness at every iteration.

In addition, we sweep a generalized positional scoring family parameterized by
exponent $p$: rank~$k$ (out of $n$) receives score $((n\!-\!1\!-\!k)/(n\!-\!1))^p$.
This family includes antiplurality ($p \to 0$), Borda ($p=1$), and plurality
($p \to \infty$) as special cases. We evaluate 28 exponents spanning
$p \in [0.05, 10]$ across the full $(K, \alpha)$ grid with 1000 runs per
configuration, identifying $p = 0.35$ as the exponent that dominates at
low-to-moderate complexity. This ninth method is included in all subsequent
analyses.

\subsection{Experiment 2: Representative Democracy}

\label{sec:exp2}

Using the representative democracy model described in Section~3, we sweep
$\beta \in \{0,\, 0.5,\, 1\}$ and $p_{\mathrm{self}} \in \{0,\, 0.5,\, 1\}$
(9 configurations) with $C=5$ candidates selected uniformly at random, $V=4$
bits voted on per round (16 proposals), across the full landscape grid:
$K \in \{1, \ldots, 20\}$ and $\alpha \in \{0.0, 0.1, \ldots, 1.0\}$.
All nine voting methods are tested, yielding 17{,}820 configurations with 500
independent runs each. The larger proposal space ($V=4$ versus $V=2$ in
Experiment~1) ensures that candidates can hold distinct platforms and
constituency welfare can meaningfully diverge from population welfare. Using
the same $(K, \alpha)$ grid as Experiment~1 allows direct comparison of the
fitted complexity axis under representation.
