We model a population of $M$ individuals making collective decisions on an NK
fitness landscape. Each individual~$i$ is represented by a binary string
$\mathbf{x}_i \in \{0,1\}^N$, partitioned into a \emph{shared portion}
$\mathbf{s} \in \{0,1\}^{N_s}$ (identical across all individuals, representing
laws) and an \emph{individual portion} $\mathbf{t}_i \in \{0,1\}^{N_t}$ (unique
and fixed, representing personal traits), where $N_s + N_t = N$.

\subsection{Fitness Landscape}

Fitness is computed according to the standard NK
model~\citep{kauffman1993origins}. Each bit position $j$ contributes a fitness
value $f_j$ that depends on the bit at position~$j$ and $K$ other positions
determined by a dependency matrix~$D$:
\begin{equation}
  F(\mathbf{x}_i) = \sum_{j=1}^{N} f_j\!\left(\mathbf{x}_i[j],\;
    \mathbf{x}_i[d_{j,1}],\ldots,\mathbf{x}_i[d_{j,K}]\right),
\end{equation}
where $d_{j,1},\ldots,d_{j,K}$ are the $K$ positions on which position~$j$
depends. Each $f_j$ maps a $(K\!+\!1)$-bit substring to a fitness value drawn
uniformly from $[-1, 1]$. Increasing $K$ increases the ruggedness of the
landscape, creating more local optima and making optimization harder.

\subsection{Cross-Dependency Parameter $\alpha$}

We control the coupling between shared and individual portions through a
cross-dependency fraction $\alpha \in [0,1]$. For each bit position~$j$, a
fraction~$\alpha$ of its $K$ dependencies are drawn from the \emph{other}
portion, while the remaining $1-\alpha$ are drawn from the same portion.
At $\alpha=0$, dependencies are entirely internal: laws affect all citizens
identically. At $\alpha=1$, dependencies are entirely cross-portion: individual
traits mediate all policy effects but lack complex internal structure. At
$\alpha \approx 0.5$, both are present, creating the highest effective
complexity.

\subsection{Voting Process}

At each time step, the system selects $V$ bit positions uniformly at random from
the shared portion (we use $V=2$ in direct democracy unless otherwise noted).
This defines $2^V$ possible
proposals (all binary configurations of those positions). Each individual~$i$
evaluates all proposals by computing the fitness they would obtain under each
one, producing a utility matrix $U \in \mathbb{R}^{M \times 2^V}$.

A voting method $\mathcal{V}$ then acts on this matrix to select a winning
proposal, which is applied to all individuals' shared portions. Equivalently, we
may view each method as a function
\begin{equation}
  \mathcal{V}: \mathbb{R}^{M \times 2^V} \to \{1,\ldots,2^V\},
\end{equation}
or as a two-stage map that first transforms $U$ into proposal scores and then
selects the maximiser. Under this interpretation, all methods solve the same
task from the same input, but they differ in which information from $U$ they
retain. Some use only each voter's top choice, others use full ordinal
rankings, others use cardinal magnitudes, and pairwise methods reprocess the
same profile through head-to-head comparisons. We study eight standard methods
plus one member of a generalized scoring family:

\begin{itemize}
  \item \textbf{Plurality}: Each voter casts one vote for their highest-utility
    proposal; the proposal with the most votes wins.
  \item \textbf{Approval}: Each voter approves all proposals that improve their
    fitness over the status quo (or all tied-for-best if none improve); the most
    approved proposal wins.
  \item \textbf{Score}: Sum each voter's raw utility for each proposal; highest
    total wins.
  \item \textbf{Borda}: Each voter ranks proposals; points are assigned by rank
    and summed.
  \item \textbf{Instant-Runoff (IRV)}: Iteratively eliminate the proposal with
    the fewest first-place votes, redistributing those votes, until a majority
    winner emerges.
  \item \textbf{STAR}: Sum scores to find the top two proposals, then elect the
    one preferred head-to-head by more voters.
  \item \textbf{Minimax (Condorcet)}: Select the proposal whose worst pairwise
    margin of defeat is minimized.
  \item \textbf{Random Dictator}: A uniformly random voter's top choice is
    selected. Serves as a baseline.
  \item \textbf{Scoring ($p\!=\!0.35$)}: A positional scoring rule where rank~$k$
    (out of $n$) receives score $((n\!-\!1\!-\!k)/(n\!-\!1))^{0.35}$. This is a
    member of the power-law scoring family that interpolates between
    antiplurality ($p \to 0$) and plurality ($p \to \infty$), with Borda at
    $p=1$. At $p=0.35$, the rule is flatter than Borda, assigning substantial
    weight across the ranking while remaining purely ordinal. This exponent was
    identified from a sweep over the scoring family as the value that dominates
    at low-to-moderate complexity.
\end{itemize}

\subsection{Representative Democracy}

In the representative variant, $C$ candidates are drawn uniformly at random
from the population. Each candidate $c$ represents a \emph{constituency}:
voters assigned by nearest-neighbor Hamming distance on individual traits.

\paragraph{Platform formation ($p_{\mathrm{self}}$).}
With probability $p_{\mathrm{self}}$, candidate $c$ acts as a \emph{trustee}
and selects the proposal maximizing their own fitness; with probability
$1 - p_{\mathrm{self}}$, they act as a \emph{delegate} and maximize mean
constituency welfare. This interpolates between the classical delegate
($p_{\mathrm{self}}=0$) and trustee ($p_{\mathrm{self}}=1$) models.

\paragraph{Candidate election ($\beta$).}
Voters evaluate candidates by a convex combination of \emph{policy utility}
$u_{\mathrm{policy}}(i,c) = F_i(\pi_c)$ and \emph{identity utility}
$u_{\mathrm{id}}(i,c) = 1 - d_H(\mathbf{t}_i, \mathbf{t}_c)/N_t$:
\begin{equation}
  u(i, c) = (1-\beta)\,\tilde{u}_{\mathrm{policy}}(i, c)
           + \beta\,u_{\mathrm{id}}(i, c),
\end{equation}
where $\beta \in [0,1]$ is the \emph{identity weight} and both components are
normalized to $[0,1]$. The same voting method used in direct democracy then
elects a winning candidate whose platform is implemented.

The $(\beta, p_{\mathrm{self}})$ plane nests four interpretable limiting cases:

\begin{center}
\begin{tabular}{lll}
\toprule
Model & $\beta$ & $p_{\mathrm{self}}$ \\
\midrule
Trustee (policy election, self-interested) & 0 & 1 \\
Delegate (policy election, constituency)   & 0 & 0 \\
Identity-trustee (identity election, self-interested) & 1 & 1 \\
Identity-delegate (identity election, constituency)   & 1 & 0 \\
\bottomrule
\end{tabular}
\end{center}

The cross-dependency $\alpha$ interacts with $\beta$: when $\alpha=0$, identity
utility carries no policy information and $\beta > 0$ introduces pure noise;
when $\alpha=1$, traits fully mediate policy effects, making identity a strong
proxy for policy alignment.
